\sc Building sex-specific synthetic lethality network in cancer - Temp Name

### 2023-08-11 sl_report_pmcrc/paper0904.tex
updates

\lhead{Synthetic Lethality} %This needs to change
\rhead{Alexander Turco}
\title{\sc Building sex-specific synthetic lethality network in cancer - Temp Name}
\author{\sc Alexander Turco}

@@

@@

\section{Next Steps/Future Avenues}
\begin{itemize}
	\item Build male and female synthetic lethal interaction networks using results from Fishers exact test. Use the genes as nodes and the odds ratios as edges to build the networks. (Packages: \texttt{NetworkX} Python, \texttt{pyvis} Python). Find differences in the male and female networks.
	\item Perform survival analysis on the SL pairs found using Fishers exact test and see which pairs contribute to improved patient survival. This could provide us with more clinically relevant synthetic lethal interactions.
	\item Incorporate other types of data. We only included RNA-seq data, what about copy number variation, methylation, etc... this could strengthen the validity of the study.
	\item Normal tissue proxy: There is a lack of normal tissue samples in the TCGA database. This restricted the number of cancer types we used in the study, as well as the sample size of the cancer types we did use. Can we use a different database to find matched samples? Can we use a machine learning approach to generate normal tissue expression data? Emailed the author of this paper, interesting stuff in terms of generating normal tissue expression but the author never got back to me with her code. \citep{zeng2019selecting}
	\item Do males have higher incidence of lung cancer because more of them are smokers, or is it because they are male. Can we try and better understand if other factors (like societal factors) cause sex differences, or is it truly due to genetic differences in sex?
	\item Menopause?
\end{itemize}

@@

This script

@@

This script

@@

This script

>>>>

\lhead{Sex Specific Synthetic Lethality} %This needs to change
\rhead{Alexander Turco}
\title{\sc Building sex-specific synthetic lethality network in cancer - Temp Name}
\author{\sc Alexander Turco}

@@

	We used the \texttt{clusterProfiler} R package to perform gene set enrichment analysis (GSEA) in order to gain insight into the functional groups of genes enriched in certain biological pathways. The package depends on annotation data from GO.db (Gene Ontology) and KEGG.db (Kyoto Encyclopedia of Genes and Genomes) in order to calculate enrichment test for GO and KEGG pathways \citep{yu2012clusterprofiler}. Essentially the analysis determines whether members of a gene set (unordered collections of genes that are functionally related) are randomly distributed in a list of differentially expressed genes, or primarily found at the top (overexpressed) or bottom (underexpressed) when sorted based on log2FoldChange \citep{yu2012clusterprofiler}.

	\begin{figure}[!h]
		\centering
		\includegraphics[width=17cm, height=10cm]{all_cancersgsea_alldegenes_male_female_bp0.9.png}
		\caption{Dotplot showing the results of GSEA performed on ALL differentially expressed genes (892 female, 1289 male) found in females (left) and males (right). This example shows 10 activated and 10 suppressed enriched gene sets.}
		\label{fig:7}
	\end{figure}

	\begin{figure}[!h]
		\centering
		\includegraphics[width=\textwidth, height=11cm]{all_cancersgsea_exclusivedegenes_male_female_bp0.9.png}
		\caption{Dotplot showing the results of GSEA performed on EXCLUSIVELY differentially expressed genes (241 female, 638 male) found in females (left) and males (right). This example shows 10 activated and 10 suppressed enriched gene sets.}
		\label{fig:8}
	\end{figure}

@@

	\begin{figure}[!h]
		\centering
		\includegraphics[width=\textwidth, height=8cm]{interactions.png}
		\caption{Barplot showing the genes involved in the most number of potential synthetic lethal pairs found using Fishers exact test}
		\label{fig:9}
	\end{figure}

\section{Next Steps/Future Avenues}
\begin{itemize}
	\item Build male and female synthetic lethal interaction networks using results from Fishers exact test. Use the genes as nodes and the odds ratios as edges to build the networks. (Packages: \texttt{NetworkX} Python, \texttt{pyvis} Python). Find differences in the male and female networks.
	\item Perform survival analysis on the SL pairs found using Fishers exact test and see which pairs contribute to improved patient survival. This could provide us with more clinically relevant synthetic lethal interactions.
	\item Incorporate other types of data. We only included RNA-seq data, what about copy number variation, methylation, etc... this could strengthen the validity of the study.
	\item Normal tissue proxy: There is a lack of normal tissue samples in the TCGA database. This restricted the number of cancer types we used in the study, as well as the sample size of the cancer types we did use. Can we use a different database to find matched samples? Can we use a machine learning approach to generate normal tissue expression data? Emailed the author of this paper, interesting stuff in terms of generating normal tissue expression but the author never got back to me with her code. \citep{zeng2019selecting}
	\item Do males have higher incidence of lung cancer because more of them are smokers, or is it because they are male. Can we try and better understand if other factors (like societal factors) cause sex differences, or is it truly due to genetic differences in sex?
	\item Menopause? Age Differences? Differences in medications?
\end{itemize}

@@

This script performs a lot of tasks.

@@

This script performs gene set enrichment analysis using \texttt{clusterProfiler}.

@@

This script performs a non parametric multivariate analysis of variance (NPMANOVA) in order to determine if differences between
